---
title: "HW 02 - Basics of Data Frames"
author: "Xander Paul"
due date: 'October 2, 2017'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
chooseCRANmirror(ind = 50)
require(readr)
install.packages("readr")
library(readr)
install.packages('ggplot2')
library(ggplot2)
install.packages('dplyr')
library(dplyr)
```

```{r}
github <- "https://github.com/ucb-stat133/stat133-fall-2017/raw/master/"
file <- "data/nba2017-player-statistics.csv"
csv <- paste0(github, file)
download.file(url = csv, destfile = 'nba2017-player-statistics.csv')
```

```{r}
base <- read.csv("nba2017-player-statistics.csv", header = TRUE, colClasses = c('character', 'character', 'factor', 'character', 'double', rep('integer', 19) ))
str(base)
```
```{r}
readr <- read_csv("nba2017-player-statistics.csv", col_types = list(Player = 'c', Team = 'c', Position = col_factor(c('C', 'PF', 'PG', 'SF', 'SG')), Experience = 'c', Salary = 'n', Rank = 'i', Age = 'i', GP = 'i', GS = 'i', MIN = 'i', FGM = 'i', FGA = 'i', Points3 = 'i', Points3_atts = 'i', Points2 = 'i', Points2_atts = 'i', FTM = 'i', FTA = 'i', OREB = 'i', DREB = 'i', AST = 'i', STL = 'i', BLK = 'i', TO = 'i'))
str(readr)
```
```{r}
experience <- replace(base$experience, base$experience=="R", 0)
experience <- as.integer(base$experience)
experience
```
```{r}
base$Experience[base$Experience=='R'] <- 0
readr$Experience[readr$Experience=='R'] <- 0
base$Experience <- as.integer(base$Experience)
readr$Experience <- as.integer(readr$Experience)
```

```{r}
Missed_FG <- base$FGA - base$FGM
base <- mutate(base, Missed_FG = Missed_FG)
Missed_FT <- base$FTA - base$FTM
base <- mutate(base, Missed_FT = Missed_FT)
pts <- 2 * base$Points2 + 3 * base$Points3 + base$FTM
base <- mutate(base, PTS = pts)
reb <- base$OREB + base$DREB
base <- mutate(base, REB = reb)
mpg <- base$MIN / base$GP
base <- mutate(base, MPG = mpg)
eff <- (pts + reb + base$AST + base$STL + base$BLK - Missed_FG - Missed_FT - base$TO) / base$GP
base <- mutate(base, EFF = eff)
summary(eff)
ggplot(data = base, aes(x = EFF)) + geom_histogram(binwidth = 5)
```

```{r}
base_some <- select(base, Player, Team, Salary, EFF)
base_some <- arrange(base_some, desc(EFF))
slice(base_some, 1:10)
```
```{r}
low_guys <- filter(base, EFF < 0)
select(low_guys, Player)
```
```{r}
cor(eff, base[c('PTS', 'REB', 'AST', 'STL', 'BLK', 'Missed_FG', 'Missed_FT', 'TO')])
cor_vector <- c(0.8588644, 0.7634501, 0.6957286, 0.6689232, 0.5679571, -0.7271456, -0.7722477, -0.8003289)
barplot(cor_vector, main = "Correlations between Player Stats and EFF", names.arg = c('PTS', 'REB', 'STL', 'AST', 'BLK', 'Missed_FT', 'Missed_FG', 'TO'), cex.names = 0.6, abline(a = 0, b = 0), col = ifelse(cor_vector > 0, 'gray', 'red'))
```

```{r}
ggplot(data = base, aes(x = EFF, y = Salary)) + geom_point() + geom_smooth(method = 'loess')
cor(eff, base[c('Salary')])
```
```{r}
players2 <- filter(base, MPG > 20)
ggplot(data = players2, aes(x = EFF, y = Salary)) + geom_point() + geom_smooth(method = 'loess')
cor(players2[c('EFF')], players2[c('Salary')])
```
Between more established players, EFF is not as strong a predictor of salary. That is, EFF is more important for determining the salary of a rookie.

# Comments

Importing the csv file was difficult and required a lot of internet searching. Downloading the packages and installing them proved to be a headache as well. 

I looked up many different R guides on the internet to help me with this assignment. The homework took me about 3 hours to complete. The most time consuming part was ensuring that the packages installed correctly, and this was highly frustrating. 

