---
title: "HW 03 - Ranking NBA Teams"
author: "Xander Paul"
due date: "October 15, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
chooseCRANmirror(ind = 50)
require(readr)
install.packages("readr")
library(readr)
install.packages('ggplot2')
library(ggplot2)
install.packages('dplyr')
library(dplyr)
```

```{r}
github <- "https://github.com/ucb-stat133/stat133-fall-2017/raw/master/"
file <- "data/nba2017-roster.csv"
csv <- paste0(github, file)
download.file(url = csv, destfile = 'nba2017-roster.csv')
roster <- read.csv('nba2017-roster.csv')

github <- "https://github.com/ucb-stat133/stat133-fall-2017/raw/master/"
file <- "data/nba2017-stats.csv"
csv <- paste0(github, file)
download.file(url = csv, destfile = 'nba2017-stats.csv')
stats <- read.csv('nba2017-stats.csv')

stats <- mutate(stats, missed_fg = stats$field_goals_atts - stats$field_goals_made)
stats <- mutate(stats, missed_ft = stats$points1_atts - stats$points1_made)
stats <- mutate(stats, points = stats$points1_made + stats$points2_made * 2 + stats$points3_made * 3)
stats <- mutate(stats, rebounds = stats$off_rebounds + stats$def_rebounds)
stats <- mutate(stats, efficiency = (stats$points + stats$rebounds + stats$assists + stats$steals + stats$blocks - stats$missed_fg -stats$missed_ft - stats$turnovers) / stats$games_played)

team <- merge(roster, stats)
teams <- summarise(group_by(team, team),
          experience = sum(experience),
          salary = round((sum(salary) / 1000000), 2),
          points3 = sum(points3_made),
          points2 = sum(points2_made),
          free_throws = sum(points1_made),
          points = sum(points),
          off_rebounds = sum(off_rebounds),
          def_rebounds = sum(def_rebounds),
          assists = sum(assists),
          steals = sum(steals),
          blocks = sum(blocks),
          turnovers = sum(turnovers),
          fouls = sum(fouls),
          efficiency = sum(efficiency))
```


```{r}
ggplot(teams, aes(x = reorder(team, salary), y = salary)) + geom_bar(stat='identity') + geom_hline(aes(yintercept = mean(salary), color = 'red')) + coord_flip() + ggtitle("NBA Teams Ranked by Total Salary")
```

```{r}
ggplot(teams, aes(x = reorder(team, points), y = points)) + geom_bar(stat='identity') + geom_hline(aes(yintercept = mean(points), color = 'red')) + coord_flip() + ggtitle("NBA Teams Ranked by Total Points")
```

```{r}
ggplot(teams, aes(x = reorder(team, efficiency), y = efficiency)) + geom_bar(stat='identity') + geom_hline(aes(yintercept = mean(efficiency), color = 'red')) + coord_flip() + ggtitle("NBA Teams Ranked by Total Efficiency")
```

## Cleveland and the Golden State Warriors tend to dominate the rankings.

```{r}
some_teams <- select(teams, points3, points2, free_throws, off_rebounds, def_rebounds, assists, steals, blocks, turnovers, fouls)
pca <- prcomp(some_teams, scale. = TRUE)
pca
```

```{r}
prop = round((pca$sdev^2), 3) / 10
cumprop = cumsum(prop)
eigs <- data.frame(eigenvalue = round((pca$sdev^2), 4), prop = prop, cumprop = cumprop)
eigs
```

```{r}
pca1 <- pca$x[ ,1]
pca2 <- pca$x[ ,2]
pctable = data.frame(pc1 = pca1, pc2 = pca2, team = teams$team)
ggplot(pctable, aes(x = pca1, y = pca2, label = team)) + geom_label() + geom_point() + ggtitle("PCA Plot (PC1 and PC2)")
```

```{r}
scaled_pc1 <- 100 * (pca1 - min(pca1)) / (max(pca1) - min(pca1))
scaled_pc1
```

```{r}
scaled <- data.frame(team = teams$team, pc1 = scaled_pc1)
ggplot(scaled, aes(x = reorder(team, pc1), y = pc1)) + geom_bar(stat='identity') + geom_hline(aes(yintercept = mean(pc1), color = 'red')) + coord_flip() + ggtitle("NBA Teams Ranked by Scaled PC1")
```

# Comments
# This was the first time I worked with this file structure, but it was fairly easy to work out. Working with relative paths was difficult considering I use Windows, but overall it worked fine. Writing code in R script is a fresh experience for me, but exciting. Doing PCA was somewhat tricky, but easy enough once I realized what the output should have been. Creating the charts was fairly simple as well. It took me about 3 hours to complete the assignment on my own. The most interesting part was doing the PCA.